NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-03-31 paper/background.tex
Minor changes (mainly in language)

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference (NLI)~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The goal of NLI is to predict the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence.
The prediction falls in one of three categories: \emph{entailment} (\textbf{E}), \emph{contradiction} (\textbf{C}), and \emph{neutral} (\textbf{N}).
A simple example is shown in Table~\ref{tab:NLI}.
In this case, the premise is ``Boy ate an apple''.
The hypothesis statement ``Kid ate fruits'' can be concluded from the premise. Therefore, the relationship between the premise and hypothesis is \emph{entailment}. However, we should note that such a relationship is not necessarily reversible. Since the concept ``fruit'' is less restrictive than that of ``apple'', we cannot conclude ``Boy ate an apple'' from the statement ``Kid ate fruit''. The same logic applies to the hypothesis of ``Boy ate a Fuji apple.'', in which the premise neither implies nor opposes the hypothesis, and therefore, their relationship is \emph{neutral}.
%\emph{Contradiction} means the textual of hypothesis opposes that of premise;
%and \emph{Neutral} implies no conclusion can be made on either \emph{Entailment} or \emph{Contradiction}.

\begin{table}[htbp]
\vspace{-2mm}
\label{tab:NLI}
\centering
\caption{An illustration of natural language inference.}
 \begin{tabular}{c | c c c c}
 \hline
  P / H & sentences & entail & contradict & neutral \\ [0.5ex]
 \hline
 premise & Boy ate an apple. &  -  &  -  & - \\
 hypothesis & Kid ate fruits. & \checkmark &   &  \\
 hypothesis & Boy ate a banana. &  & \checkmark & \\
 hypothesis & Tom ate an apple. &  &  & \checkmark \\
 hypothesis & Boy ate a Fuji apple. &   &  & \checkmark \\
 %\hline
 %premise & Facebook's IPO electrified the general public. &  -  &  -  & - \\
 %hypothesis & Facebook went public. & \checkmark   &  &  \\
 %hypothesis & General Electric went public. &  &   & \checkmark \\
 %hypothesis & People ignored Facebook's IPO. &  & \checkmark & \\
 \hline
\end{tabular}
\vspace{-2mm}
\end{table}

At first glance, the task of natural language inference may seem less practical compared to other well-known NLP challenges such as machine translation; however, the ability to distinguish entailment and contradiction relationship is fundamental to understanding natural language at large.
%
Considering the ambiguousness of natural language and polysemy words, the inference task can become quite challenging (especially from the learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refers to premise, \textbf{H} refers to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal similarity between ``Electric'' and ``electrified'' may trick a model to predict \textbf{H2} as \emph{entailment}. A model likely will also fail to understand the link between ``went public'' and ``IPO'', and therefore, mistake \textbf{H1} as \emph{neutral}.
%
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduced a large corpus~\cite{BowmanAngeliPotts2015} for NLI tasks, which has helped spawn a new wave of effective neural network models for those tasks. Here, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset (see \textbf{Appendix A} for a detailed description of the model).

@@

\subsection{Neural Network Model in NLP}
%\shusen{What is the basic understanding of neural network What is the end-to-end}
There are many connections and distinctions between neural network models employed for vision and NLP tasks.
On the one hand, as discussed in the introduction, the discrete nature of words and sentences presents additional challenges for
interpreting the NLP model.
%
On the other hand, a majority of recent NLP neural networks share the nature of
end-to-end models, where the entire model operates as a black-box that takes
vectorized input and yields a final prediction for a specific task.
%

Many of the recent NLP end-to-end neural network models, despite having potentially drastically different network architectures, share a similar high-level design that consists of three distinct stages (encode, align, classify, see Fig.~\ref{fig:end2end}).
%
The existence of shared conceptual structures means the proposed tool, although designed for NLI tasks, can be readily generalized for a much broader set of applications.

@@

\subsection{Attention Mechanism}
\label{sec:attention}
Among the three stages, the second stage often constitutes a crucial part of the model as it determines where the classifier will focus for generating a prediction. The operation to compute the alignment between words in the input is referred to as the attention mechanism. 
The introduction of the attention mechanism~\cite{bahdanau2014neural} allows
pairwise interaction between internal representations of words. 
This interaction can be naturally explained as a form of alignment that exposes an interpretable layer in end-to-end neural networks.
%
%As the only interpretable part of the internal state of the network, understanding how attention information 
Recently, attention has contributed to many strongly performed NLP models~\cite{parikh2016emnlp,rush2015neural,yang2016hierarchical,seo2016bidirectional,schwartz2017high}.

>>>>

\subsection{Natural Language Inference}
\label{sec:languageInference}
Natural Language Inference (NLI)~\cite{DaganRothSammons2013} is an important machine understanding task in NLP.
The goal of NLI is to predict the relationship between a premise (\textbf{P}) sentence and a hypothesis (\textbf{H}) sentence.
The prediction falls in one of three categories: \emph{entailment} (\textbf{E}), \emph{contradiction} (\textbf{C}), and \emph{neutral} (\textbf{N}).
A simple example is shown in Table~\ref{tab:NLI}.
In this case, the premise is ``A boy ate an apple''.
The hypothesis statement ``A kid ate fruit'' can be concluded from the premise. Therefore, the relationship between the premise and hypothesis is \emph{entailment}. However, we should note that such a relationship is not necessarily reversible. Since the concept ``fruit'' is less restrictive than that of ``apple'', we cannot conclude ``A boy ate an apple'' from the statement ``A kid ate fruit''. The same logic applies to the hypothesis of ``A boy ate a Fuji apple.'', in which the premise neither implies nor opposes the hypothesis, and therefore, their relationship is \emph{neutral}.
%\emph{Contradiction} means the textual of hypothesis opposes that of premise;
%and \emph{Neutral} implies no conclusion can be made on either \emph{Entailment} or \emph{Contradiction}.

\begin{table}[htbp]
\vspace{-2mm}
\label{tab:NLI}
\centering
\caption{An illustration of natural language inference.}
 \begin{tabular}{c | c c c c}
 \hline
  P / H & sentences & entail & contradict & neutral \\ [0.5ex]
 \hline
 premise & A boy ate an apple. &  -  &  -  & - \\
 hypothesis & A kid ate fruit. & \checkmark &   &  \\
 hypothesis & A boy ate a banana. &  & \checkmark & \\
 hypothesis & Tom ate an apple. &  &  & \checkmark \\
 hypothesis & A boy ate a Fuji apple. &   &  & \checkmark \\
 %\hline
 %premise & Facebook's IPO electrified the general public. &  -  &  -  & - \\
 %hypothesis & Facebook went public. & \checkmark   &  &  \\
 %hypothesis & General Electric went public. &  &   & \checkmark \\
 %hypothesis & People ignored Facebook's IPO. &  & \checkmark & \\
 \hline
\end{tabular}
\vspace{-2mm}
\end{table}

At first glance, the task of natural language inference may seem less practical compared to other well-known NLP challenges such as machine translation; however, the ability to distinguish entailment and contradiction relationship is fundamental to understanding natural language at large.
%
Considering the ambiguousness of natural language and polysemy of words, the inference task can become quite challenging (especially from the learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refers to premise, \textbf{H} refers to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal similarity between ``Electric'' and ``electrified'' may trick a model to predict \textbf{H2} as \emph{entailment}. A model likely will also fail to understand the link between ``went public'' and ``IPO'', and therefore, mistake \textbf{H1} as \emph{neutral}.
%
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduced a large corpus~\cite{BowmanAngeliPotts2015} for NLI tasks, which has helped spawn a new wave of effective neural network models for those tasks. Here, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset (see \textbf{Appendix A} for a detailed description of the model).

@@

\subsection{Neural Network Models in NLP}
%\shusen{What is the basic understanding of neural network What is the end-to-end}
Neural network models employed for computer vision and NLP tasks share a key common characteristic: a majority of recent approaches use end-to-end models. That is, the entire model operates as a black-box that takes vectorized inputs and yields a final prediction for a task of interest.
However, as discussed in the introduction, the discrete nature of words and sentences presents additional challenges for
interpreting NLP models.
%

Many recent end-to-end neural network models in NLP, despite having  diverse network architectures, share a similar high-level design that consists of three distinct stages (encode, align, classify, see Fig.~\ref{fig:end2end}).
%
The existence of shared conceptual structures means the proposed tool, although designed for NLI tasks, can be readily generalized for a much broader set of applications.

@@

\subsection{Attention Mechanism}
\label{sec:attention}
Among the three stages, the second stage often constitutes a crucial part of the model as it determines where the classifier will focus for generating a prediction. The operation to compute the alignment between words in the input is referred to as the attention mechanism~\cite{bahdanau2014neural}. 
The introduction of the attention mechanism allows
pairwise interaction between internal representations of words. 
This interaction can be naturally explained as a form of alignment that exposes an interpretable layer in end-to-end neural networks.
%
%As the only interpretable part of the internal state of the network, understanding how attention information 
Recently, attention has contributed to the strong performance of many NLP models~\cite{parikh2016emnlp,rush2015neural,yang2016hierarchical,seo2016bidirectional,schwartz2017high}.

### 2018-03-31 paper/intro.tex
Minor changes (mainly in language)

We implement our vision of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
However, the components of the visualization and the overall paradigm can be readily extended to other NLP tasks, such as question and answer, text summarization, etc.
%
In the simplest form, the inference task asks the question about whether \textbf{sentence A} and \textbf{sentence B}: (1) are \emph{entailed} (one can infer \textbf{B} from \textbf{A}; (2) \emph{contradict} (\textbf{B} disagrees with \textbf{A}) or (3) are \emph{neutral} (\textbf{A} and \textbf{B} talk about different/unrelated things
% we can infer the content of \textbf{sentence B} from \textbf{sentence A} (\emph{entail}), if they contradict with each other (\emph{contradiction}), or whether two sentences talk about different/unrelated things (\emph{neutral}). 
Language inference addresses the fundamental challenge of measuring the semantic similarity of sentences, and is one of the core NLP tasks (see Section~\ref{sec:languageInference} for details).

One recent advance in neural natural language process models is the introduction of \emph{attention mechanisms}~\cite{VaswaniShazeerParmar2017} (Section~\ref{sec:attention}). Intuitively, the attention helps capture which part of the input the model is deemed more important for making a prediction. The attention is often represented as weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many speculations on how attention works in various models. By utilizing the proposed tool, the domain experts study how changes in sentence input, attention, or prediction affect each other, which helps them develop deeper intuitive and alternative hypotheses. 
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label. 
%Another word, 
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
Moreover, we propose to enhance the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar-guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:grammarAttention}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate about the potential benefits of including the linguistic structure in the design of the attention component of the model.

In summary, the key contributions of the proposed works are the following:
\begin{itemize}
  % \item Distill the visualization tasks for making sense of the role of attention in natural language inference (section~\ref{sec:task});
    \item A perturbation-driven exploration paradigm derived from close examination of how domain experts conduct exploratory analysis on end-to-end NLP neural network models;

    \item The NLIZE system that enriches such a paradigm by providing an intuitive environment that allows domain experts to readily express hypotheses and obtain instantaneous feedbacks; 
    %to  experiment with the intricate relationships among input sentences, model internals, and the prediction;
    %\item Perform error analysis by examining the stability of the predictions under perturbation;

    \item An optimization scheme for correcting a failed prediction based on the natural extension of the margin-infused relaxed algorithm (MIRA) to the neural network; and
    % to minimally update the model to fix a failed prediction, which help reason about the  better understand the re the question about how t

>>>>

We implement our vision of the perturbation-driven exploration in an interactive visualization system for natural language inference models~\cite{Parikh2016}.
However, the components of the visualization and the overall paradigm can be readily extended to other NLP tasks, such as question and answer, text summarization, etc.
%
In the simplest form, the inference task asks whether the
relationship between \textbf{sentence A} and \textbf{sentence B}: (1) is \emph{entailment} (one can infer \textbf{B} from \textbf{A}), (2) \emph{contradiction} (\textbf{B} disagrees with \textbf{A}) or (3) is \emph{neutral} (\textbf{A} and \textbf{B} talk about different/unrelated things).
% we can infer the content of \textbf{sentence B} from \textbf{sentence A} (\emph{entail}), if they contradict with each other (\emph{contradiction}), or whether two sentences talk about different/unrelated things (\emph{neutral}). 
Natural language inference addresses the fundamental challenge of identifying semantic relationships between sentences, and is a core NLP task (see Section~\ref{sec:languageInference} for details).

One recent advance in neural natural language process models is the
introduction of \emph{attention
  mechanisms}~\cite[for example]{bahdanau2014neural,VaswaniShazeerParmar2017}
(Section~\ref{sec:attention}). Intuitively, attention asks which parts of the
input are deemed more important for making a prediction. The attention is often represented via weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many speculations on how attention works in various models.
Using the proposed tool, the domain experts can study how changes in sentence input, attention, or prediction affect each other, which helps them develop deeper intuitive and alternative hypotheses. 
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label. 
%Another word, 
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
Moreover, we propose to enhance the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar-guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:grammarAttention}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate about the potential benefits of including the linguistic structure in the design of the attention component of the model.

In summary, the key contributions of this paper are:
\begin{itemize}
  % \item Distill the visualization tasks for making sense of the role of attention in natural language inference (section~\ref{sec:task});
    \item A perturbation-driven exploration paradigm derived from close examination of how domain experts conduct exploratory analysis on end-to-end NLP neural network models;

    \item The NLIZE system that enriches such a paradigm by providing an intuitive environment that allows domain experts to readily express hypotheses and obtain instantaneous feedback; 
    %to  experiment with the intricate relationships among input sentences, model internals, and the prediction;
    %\item Perform error analysis by examining the stability of the predictions under perturbation;

    \item An optimization scheme for correcting a failed prediction based on a natural extension of the margin-infused relaxed algorithm (MIRA) to neural networks; and
    % to minimally update the model to fix a failed prediction, which help reason about the  better understand the re the question about how t

### 2018-03-31 paper/task.tex
Minor changes (mainly in language)

During our long-term collaboration with NLP experts, we have conducted extensive discussions to understand the common approach employed by researcher for assessing the behavior of a model.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.

>>>>

During our long-term collaboration with NLP experts, we have conducted extensive discussions to understand the common approach employed by researchers for assessing the behavior of a model.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.
